---
title: 'R Tutorial #1 -- Econ 103'
output:
  html_document:
    css: custom.css
    theme: readable
    toc: yes
  pdf_document:
    highlight: zenburn
    toc: yes
---
Part 1: Installing R
--------------------------------------------------------------------------------
Carry out the following two steps *in order*

1. Go to http://cran.r-project.org/ and install the version of R for your 
operating system.

2. Go to http://rstudio.org/download/desktop and click the link listed under 
"Recommended for Your System". Follow the instructions to install RStudio.

To make sure this worked, open the program RStudio and go to 
File > New > R Script (or Ctrl + Shift + N). In the document, type the text 
given in the box below, highlight both lines of code, and click "Run" 
(or Ctrl + Shift + S). If everything is working correctly, the console should 
display ``"Hello World!"`` and ``TRUE``.
```{r results='hide'}
print("Hello World!")
x <- 5
x == 5
```

Congratulations on writing your first R script! To save it, go to 
File > Save As, and choose a name. NOTE: Always save your scripts as .R files 
so they'll open in RStudio by default. 

Note that you can run one line of your script at a time by moving your cursor 
to that line and pressing CONTROL-ENTER or COMMAND-RETURN depending on whether 
you're running Mac OSX, Linux or Windows. Another helpful shortcut 
is CONTROL-A  (COMMAND-A on Mac) which highlights all of the lines of code 
in the text editor.


Part 2: Try R Tutorial
--------------------------------------------------------

In this section, you'll be completing levels 1-6 on Try R, an online tutorial 
for learning R.

1. Go to http://www.codeschool.com/courses/try-r

2. Complete levels 1-6. After you complete each level, go back to RStudio and 
solve the corresponding exercises given in Part 4 below to test your 
understanding. Save your work in an R script. 

For your reference, I have posted a complete transcript of the commands 
introduced in Try R on the course webpage.


Part 3: More Commands
--------------------------------------------------------
Try R gives a good foundation for starting out with R, but there are some 
important commands it doesn't mention. Once you've finished Try R, work through 
this material. Corresponding exercises are given in Part 4 below. You may also 
want to read this part before attempting the exercises for Try R Section 6.

### Coding Style
Just like in writing, there are certain style rules that help improve 
readability such as using punctuation, spaces, and capitalization (looking at 
you, E E Cummings!). While coding style differs greatly across languages and
across groups, the key is to remain consistent so that your code is easily 
readable. If you have not developed a coding style yet, I would recommend 
following Google's R Style Guide (https://google.github.io/styleguide/Rguide.xml)
and adapt it as necessary. Remember: BE CONSISTENT

### Variable Names
You can use nearly anything as a variable name in R. The only rules are:

  1. Must not start with a number
  2. Must not include any characters that have a "special meaning" such 
  as ``*`` ``+`` and so on

In practical terms, this means you should only use letters, numbers, the 
underscore ``_`` and periods ``.`` in your variable names. It's also good 
practice not to give variables the same names as existing R commands. 
For example ``mean`` would not be a good choice for a variable name 
but ``sampleMean`` is fine.

[A note for those of you who have programming experience: while R supports 
object-oriented programming, periods ``.`` do *not* have a special meaning in 
the language. For historical reasons, R programmers often use periods in place 
of underscores in variable names, but either works. Just be consistent to 
keep your code readable.]

### Documentation
Google will be your best friend in learning how to code. If you know a command,
but don't remember how it works, you can either use ``help(command)`` 
or ``?command`` to pull up the documentation for the specified command.

### Data Frames and Tables
Try R introduced the concept of data frames to store data. While the "data.frame"
package can accomplish everything we are going to do in this course, the 
"data.table" package is optimized for larger data sets (see https://github.com/Rdatatable/data.table/wiki/Benchmarks-:-Grouping for some 
benchmarks comparing "data.table" to "data.frame"). Below is a short intro
to the "data.table" syntax so that you will be equipped to handle larger datasets
in the future (for a more in-depth tutorial, check out https://rawgit.com/wiki/Rdatatable/data.table/vignettes/datatable-intro.html). 

We'll be using the salary data for Chicago city employees to introduce us 
to the "data.table" syntax.

#### Install ``data.table``
``install.packages("data.table")``

#### Reading Data
Instead of using the ``read.csv`` or ``read.table`` commands, use the ``fread``
(it stands for "fast read") command to load any structured data set (doesn't 
matter if it is a .csv, tab-delimited, or other commonly delimited files)
```{r}
library(data.table)
salaryData <- fread("https://data.cityofchicago.org/api/views/xzkq-xp2w/rows.csv?accessType=DOWNLOAD")
```

We can rename the columns as we desire (also, shorter names are easier to type).
```{r}
names(salaryData) <- c("Name", "Position", "Dept", "Salary")
head(salaryData)
```

#### Subsetting Data
Let's say that we're only interested in looking at the police department. 
I'll make a subset of only the data related to the police department.
To get the corresponding subset, we need to figure out 
which column is the department (it's "Dept", since I changed it). 
The below command only returns the entries for which
the condition is true ("==" is a logical operator).
```{r}
copSalaries <- salaryData[Dept == "POLICE"]
head(copSalaries)
```

#### Exploring Data
Unfortunately, since the salary data has "$", R thinks it is a character string
instead of numbers. To get rid of that, we need to strip out the dollar signs 
and convert it to numeric data (slightly advanced, so it's okay if you don't 
completely follow what's happening here).
```{r}
copSalaries$Salary <- gsub('\\$', '', copSalaries$Salary)
copSalaries$Salary <- as.numeric(copSalaries$Salary)
```

Let's sort the data by salary since that's usually the first thing anyone 
would do. We have to insert the "-" to sort from highest to lowest, otherwise
it defaults to sorting from lowest to highest.
```{r}
copSalaries <- copSalaries[order(-Salary)]
```

The top 5 highest paid people in the CPD are as follows:
```{r}
head(copSalaries, n = 5)
```

No surprise there, the police chiefs and superintendent are the highest paid in
the CPD.

How many people make over \$100K? To do this easily, we tell R to sum the number
of salary observations that are greater than \$100K.
```{r}
over100K <- copSalaries[, sum(Salary > 100000)]
print(over100K)
```
Wow! Over 1700 people in the CPD make over \$100K!

### Making Data Tables and Matrices
While you will generally be loading data into R for analysis instead of creating
it yourself, there are times where you will want to create small datasets 
to test your code before you run it on a 1GB (or larger) data file.

To make a data table, it's easiest to create the columns that will populate your
data table first and then combine them together.

```{r}
person <- c("Linus", "Snoopy", "Lucy", "Woodstock")
age <- c(5, 8, 6, 2)
weight <- c(40, 25, 50, 1)
myData <- data.table(person, age, weight)
myData
```

Part 4: Exercises
--------------------------------------------------------
### Basics

1. Calculate how many minutes there are in a week.
```{r echo=FALSE,results='hide',fig.show='hide'}
60 * 24 * 7
```
2. Add up the numbers 1 8 4 2 9 4 8 5 *without using any plus signs*
```{r echo=FALSE,results='hide',fig.show='hide'}
sum(1, 8, 4, 2, 9, 4, 8, 5)
```
3. You've forgotten what the function ``rep`` does. Load the corresponding help file.
```{r echo=FALSE,results='hide',fig.show='hide'}
help(rep)
?rep
```
4. Suppose I ran the following R commands in order. What result would I get? 
*Do not* use R to answer this: think it through and then check your answer.
  - ``x <- 5``
  - ``y <- 7``
  - ``z <- x + y``
  - ``z + 3 == 15``
```{r echo=FALSE,results='hide',fig.show='hide'}
x <- 5
y <- 7
z <- x + y
z + 3 == 15
```
5. How can I get R to print out "Go Penn!" thirty times without repeatedly 
typing this by hand?
```{r echo=FALSE,results='hide',fig.show='hide'}
rep("Go Penn!", times = 30)
```
6. Create a vector called ``x`` containing the sequence -1, -0.9, ... 0, 0.1, ..., 0.9, 1 
and then display the result
```{r echo=FALSE,results='hide',fig.show='hide'}
x <- seq(-1, 1, by = 0.1)
x
```

### Vectors (Harry Potter Style)
1. Create two vectors: ``wizards`` and ``ranking``. The vector ``wizards`` 
should contain the following names: Harry, Ron, Fred, George, Sirius. The 
vector ``ranking`` should contain the following numbers: 4, 2, 5, 1, 3 in it. 
```{r echo=FALSE,results='hide',fig.show='hide'}
wizards <- c("Harry", "Ron", "Fred", "George", "Sirius")
ranking <- c(4, 2, 5, 1, 3)
# Remember that the elements of character vectors need to be enclosed in 
# quotation marks. Either single or double quotes will work.
```
3. Extract the second element of the vector ``wizards``.
```{r echo=FALSE,results='hide',fig.show='hide'}
wizards[2]
```
4. Replace the names Fred, George and Sirius in the vector wizards with 
Hermione, Ginny, and Malfoy, respectively.
```{r echo=FALSE,results='hide',fig.show='hide'}
#There are several different ways to do this. Here are two possibilities.
wizards[c(3, 4, 5)] <- c("Hermione", "Ginny", "Malfoy")
wizards[3:5] <- c("Hermione", "Ginny", "Malfoy")
```
5. Someone who hasn't read Harry Potter needs labels to determine who these 
characters are. Assign names to the elements of the vector ``wizards``: 
Lead, Friend, Friend, Wife, Rival. Display the result.
```{r echo=FALSE,results='hide',fig.show='hide'}
names(wizards) <- c("Lead", "Friend", "Friend", "Wife", "Rival")
wizards
```
6. An avid reader of Harry Potter argues that Malfoy is not Harry's rival by 
the end of the series. Change Rival to Ex-Rival.
```{r echo=FALSE,results='hide',fig.show='hide'}
names(wizards)[5] <- "Ex-Rival"
names(wizards)
```
7. Make a barplot of the vector ``ranking``. Why can't you make a barplot of the vector ``wizards``?
```{r echo=FALSE,results='hide',fig.show='hide'}
barplot(ranking)
#The wizards vector contains character data rather than numerical data so we can't plot it.
```
8. Assign the elements of the vector ``wizards`` to be the names of the vector 
``ranking``. Then re-do the barplot so we can see who's who.
```{r echo=FALSE,results='hide',fig.show='hide'}
names(ranking) <- wizards
barplot(ranking)
```

### More Vectors and Charts (Steve's Personal Finances)
1. In 2009 Steve's income was $50,000 and his total expenses were $35,000. 
In 2010 his income was $52,000 and his expenses were $34,000. 
In 2011, his income was $52,500 and his expenses were $38,000. 
Finally, in 2012 Steve's earnings were $48,000 and his expenses were $40,000. 
Create three vectors to store this information in 
parallel: ``years``, ``income`` and ``expenses``.
```{r echo=FALSE,results='hide',fig.show='hide'}
years <- c(2009, 2010, 2011, 2012)
income <- c(50000, 52000, 52500, 48000)
expenses <- c(35000, 34000, 38000, 40000)
```
2. Following on from the previous question, calculate Steve's annual savings 
and store this in a vector called ``savings``. Make a scatterplot of ``years`` 
against ``savings``.
```{r echo=FALSE,results='hide',fig.show='hide'}
savings <- income - expenses
plot(years, savings)
```
3. Assuming zero interest on bank deposits (roughly accurate at the moment), 
calculate the total amount that Steve has saved over all the years for which 
we have data.
```{r echo=FALSE,results='hide',fig.show='hide'}
sum(savings)
```
4. Redefine the vector ``years`` so that it runs from 2009-2013. 
Redefine ``income`` to match but record ``income`` for 2013 as ``NA`` 
(this is R's code for missing data). How can we compute the sum of the elements 
of this new ``income`` vector (ignoring the ``NA``)?
```{r echo=FALSE,results='hide',fig.show='hide'}
years <- c(2009, 2010, 2011, 2012, 2013)
income <- c(50000, 52000, 52500, 48000, NA)
#If we just use sum, we'll get NA
sum(income)
#To avoid this, we use the following option to ignore the NA
sum(income, na.rm = TRUE)
```

### Summary Stats

1. Twenty-six students took the midterm. Here are their scores:
18, 95, 76, 90, 84, 83, 80, 79, 63, 76, 55, 78, 90, 81, 88, 89, 92, 73, 83, 
72, 85, 66, 77, 82, 99, 87. Assign these values to a vector called ``scores``.
```{r echo=FALSE,results='hide',fig.show='hide'}
scores <- c(18, 95, 76, 90, 84, 83, 80, 79, 63, 76, 55, 78, 90, 81, 88, 89, 92, 73, 83, 72, 85, 66, 77, 82, 99, 87)
```
2. Calculate the mean, median and standard deviation of the scores.
```{r echo=FALSE,results='hide',fig.show='hide'}
mean(scores)
median(scores)
sd(scores)
```

### Data Tables (Trojan War)

1. Create three vectors. First store the numeric values 21, 26, 51, 22, 160, 
160, 160 in a vector called ``age``. Next, store the names Achilles, Hector, 
Priam, Paris, Apollo, Athena, Aphrodite in a character vector called ``person``. 
Finally store the words Aggressive, Loyal, Regal, Cowardly, Proud, Wise, 
Conniving in a vector called ``description``
```{r echo=FALSE,results='hide',fig.show='hide'}
age <- c(21, 26, 51, 22, 160, 160, 160)
person <- c("Achilles", "Hector", "Priam", "Paris", "Apollo", "Athena", "Aphrodite")
description <- c("Aggressive", "Loyal", "Regal", "Cowardly", "Proud", "Wise", "Conniving")
```
2. Create a data table called ``trojanWar`` whose columns contain the vectors 
from the previous question.
```{r echo=FALSE,results='hide',fig.show='hide'}
trojanWar <- data.table(person, age, description)
```
3. Suppose you wanted to display only the column of ``trojanWar`` that contains 
each person's ``description``. What command would you use?
```{r echo=FALSE,results='hide',fig.show='hide'}
#There are many different ways to do this:
trojanWar$description
trojanWar[["description"]]
```
4. What command would you use to show information for Achilles and Hector only?
```{r echo=FALSE,results='hide',fig.show='hide'}
#There are several ways to do this. Here are a few:
trojanWar[person == "Achilles" | person == "Hector", ]
trojanWar[person %in% c("Achilles", "Hector"), ]
trojanWar[c(1, 2), ]
trojanWar[1:2, ]
```
5. What command would you use to display the ``person`` and ``description`` columns for Apollo, Athena and Aphrodite only?
```{r echo=FALSE,results='hide',fig.show='hide'}
#There are many ways to do this. Here are a few:
trojanWar[person %in% c("Apollo", "Athena", "Aphrodite"), .(person, description)]
trojanWar[c(5, 6, 7), c(1, 3)]
trojanWar[5:7, c("person", "description")]
```
6. It turns out there was a mistake in the data: Priam's age should be 72 rather than 51. Make the appropriate change to ``trojanWar``
```{r echo=FALSE,results='hide',fig.show='hide'}
# There are many different ways to do this. The first is the most foolproof, but
# here are others:
trojanWar$age[trojanWar$person == "Priam"] <- 72
trojanWar[3, 2] <- 72
trojanWar$age[3] <- 72
trojanWar
```
