---
title: 'R Tutorial #2 -- Econ 103'
output:
  html_document:
    css: custom.css
    theme: readable
    toc: yes
  pdf_document:
    highlight: zenburn
    toc: yes
  word_document:
    toc: yes
---
This tutorial has two parts. In the first part, you'll compute summary 
statistics and graphs using R. In the second part, you'll use what you've 
learned to explore a real-world dataset: the passenger manifest from the Titanic.

## Summary Stats and Graphics

### Load and View the Data
In this section we'll work with car data that is preloaded in R. First, make sure
you have the "datasets" package loaded and then create a data.table with the data.
```{r}
library(datasets)
carData <- mtcars
```
It's always a good idea to take a look at your data after loading it. 
The functions ``head`` and ``tail``, introduced in the previous tutorial, 
make this easy (this dataset is small enough that you can see the whole thing too):
```{r}
head(carData)
tail(carData)
```

### Making Histograms
Making beautiful charts is a valuable skill. Unfortunately, the default R graphics
are really ugly. Thankfully, there's a package for that! You'll need to install 
the ``plotly`` package to generate the prettier charts. Another huge plus is that
it makes **interactive charts**!

Generating a chart is a simple matter of using the command 
``plot_ly(data, xData, yData, chartType, mode)`` (there's tons of other options, but 
we'll only focus on the basics). A histogram is just a frequency chart of one 
variable, so there's no need to specify the ``y`` variable when you're making
a histogram.
```{r}
library(plotly)
plot_ly(data = carData, x = mpg, type = "histogram")
```
The chart isn't too well labeled. Having a title and proper axis labels would 
be nice. Let's add the proper labels. 
```{r}
plot_ly(data = carData, x = mpg, type = "histogram") %>% 
  layout(title = "MPG Histogram", xaxis = list(title = "MPG"), 
         yaxis = list(title = "Frequency"))
```

By default, R produces histograms in terms of *frequencies*. That is, it counts 
the *numbers* of observations in each bin. To change this to 
*relative frequencies*, that is proportions, we add the argument ``histnorm``
and set it to ``percent``. Note, I also changed the y-axis label.
```{r}
plot_ly(data = carData, x = mpg, type = "histogram", histnorm = "percent") %>% 
  layout(title = "MPG Histogram", xaxis = list(title = "MPG"), 
         yaxis = list(title = "Percent"))
```

### Scatter Plots
Making scatter plots is very similar to how we made histograms, with the 
appropriate change to the plot ``type`` as well as specifying the ``y`` data. 
Let's make a scatter plot of horsepower and MPG. What happens if you forget to 
specify ``mode = "markers"``?
```{r}
plot_ly(data = carData, x = mpg, y = hp, type = "scatter", mode = "markers") %>% 
  layout(title = "MPG and Horsepower", xaxis = list(title = "MPG"), 
         yaxis = list(title = "Horsepower"))
```

Let's say your were interested in how different engine sizes were related to this 
scatter. You could color the dots by the number of cylinders.
```{r}
plot_ly(data = carData, x = mpg, y = hp, type = "scatter", 
        mode = "markers", color = cyl) %>% 
  layout(title = "MPG and Horsepower", xaxis = list(title = "MPG"), 
         yaxis = list(title = "Horsepower"))
```
If you're familiar with cars, the color grouping should not surprise you.

### Box Plots
A boxplot is an alternative way to visualize the distribution of a dataset. 
Once again, a few tweaks to the chart type should give us what we need.
```{r}
plot_ly(data = carData, y = mpg, type = "box", boxpoints = FALSE) %>% 
  layout(title = "MPG Box Plot")
```
Let's say you're interested in the MPG spread by engine size. How can you easily 
create the box plots for those different groups? Just tell R that you want to 
group the data by ``cyl`` and you're good to go!
```{r}
plot_ly(data = carData, y = mpg, type = "box", boxpoints = FALSE, 
        group = cyl) %>% 
  layout(title = "MPG Box Plot")
```

### Summary Statistics
The ``summary`` command takes a *whole dataframe* as its argument, unlike the other 
commands for summary statistics. It gives us the mean of each column along 
with the five-number summary. It also indicates if there are any missing 
observations (NA's) in the columns:
```{r}
summary(carData)
```

### ``mean``
Calculates the sample mean of a numeric vector
```{r}
mean(carData$mpg, na.rm = TRUE)
```

### ``median``
Calculates the sample median of a numeric vector
```{r}
median(carData$mpg, na.rm = TRUE)
```

### ``var``
Calculates the sample variance of a numeric vector
```{r}
var(carData$mpg, na.rm = TRUE)
```

### ``sd`` 
Calculates the sample standard deviation of a numeric vector
```{r}
sd(carData$mpg, na.rm = TRUE)
```
This is identical to the (positive) square root of the sample variance:
```{r}
sqrt(var(carData$mpg, na.rm = TRUE))
```

### ``IQR``
Calculates the interquartile range of a numeric vector (the 75th percentile minus the 25th percentile)
```{r}
IQR(carData$mpg, na.rm = TRUE)
```

### ``max, min``
These do exactly what they say
```{r}
max(carData$mpg, na.rm = TRUE)
min(carData$mpg, na.rm = TRUE)
```
and can be used to calculate the range
```{r}
max(carData$mpg, na.rm = TRUE) - min(carData$mpg, na.rm = TRUE)
```
To get both the maximum and minimum at once, you can use the function range;
```{r}
range(carData$mpg, na.rm = TRUE)
```
Note that this does *not* compute the *summary statistic* called the range. 

### ``quantile``
This function calculates sample quantiles, aka percentiles, of a numeric vector. 
If you simply pass it a numeric vector with no other arguments, it will give 
you the five-number summary:
```{r}
quantile(carData$mpg, na.rm = TRUE)
```
You can ask for specific quantile by using the argument probs
```{r}
quantile(carData$mpg, na.rm = TRUE, probs = 0.3)
```
Note that quantiles are specified as *probabilities* so the preceding command 
gives the 30th percentile. You can also ask for multiple quantiles at once:
```{r}
quantile(carData$mpg, na.rm = TRUE, probs = c(0.1, 0.3, 0.7, 0.9))
```

### ``cor``
This command calculates sample correlation coefficients. If you pass it two vectors of the same length it returns the correlation between them. Unlike the univariate summary statistics, in which we used ``na.rm = TRUE`` to ignore missing observations, for ``cor`` we use the argument ``use = "complete.obs"``. The reason for this difference is that ``cor`` can handle a matrix or dataframe as its input. In this case there are actually many different ways of handling missing observations. For more details, see the help file. Setting  ``use = "complete.obs"`` removes any rows with missing observations before proceeding.
```{r}
cor(carData$mpg, carData$hp, use = "complete.obs")
```
We see that there is a large negative correlation between ``mpg`` and ``hp``

You can also put a whole data table into ``cor`` and it will return the whole matrix
of all correlations between each variable (i.e. *all pairwise correlations*). 
```{r}
cor(carData, use = "complete.obs")
```
We can see that ``mpg`` is also negatively correlated with ``cyl`` (engine cylinders),
``wt`` (car weight), but it is positively correlated with ``qsec`` (the time to 
cover 0.25 miles). The preceding matrix has ones on its main diagonal since the 
correlation between any variable and itself is one. (A good exercise for extra 
practice would be to prove this assertion using the formula for correlation from 
class. It's not very difficult.)

### ``cov``
This command works just like ``cor`` but returns covariances rather than correlations:
```{r}
cov(carData$mpg, carData$hp, use = "complete.obs")
cov(carData, use = "complete.obs")
```

## Exploring the Titanic
Now it's your turn: using what you've learned above, you'll analyze some data
about everyone that was aboard the Titanic! You already know enough R to answer 
some interesting questions, so let's dive in! 

#### A Helpful Hint: Taking the mean of a vector of ones and zeros is the *same thing* as calculating the *proportion of ones*.

1. Read the documentation file for the titanic dataset so that you know what 
each column in the data represents: https://github.com/emallickhossain/Econ103Public/raw/master/Rtutorials/titanic.txt

2. Download the data, store it in a data table called titanic, and display the 
first few rows.
```{r echo=FALSE,results='hide',fig.show='hide'}
#Ex-2
library(data.table)
titanic <- fread("https://github.com/emallickhossain/Econ103Public/raw/master/Rtutorials/titanic3.csv")
head(titanic)
```

3. We'll only be using the following columns of titanic for this example: pclass, 
survived, sex, age, and fare. Select only these columns and store then in the 
data table titanic, overwriting the original dataset. Display the first few 
rows to make sure your command worked.
```{r echo=FALSE,results='hide',fig.show='hide'}
#Ex-3
titanic <- titanic[, list(pclass, survived, sex, age, fare)]
head(titanic)
```

4. Using the command summary to get an overview of the dataset and answer the following questions:
  - Are there any missing values in this dataset?
  - What was the average age of passengers on the Titanic?
  - What was the average ticket price for passengers on the ship?
  - Do the ticket prices show evidence of skewness? If so, are they positively or negatively skewed?
  - Were there more men or women aboard the Titanic?
  - What proportion of the passengers survived the disaster?
  - To which social class did most of the passengers belong?
```{r echo=FALSE,results='hide',fig.show='hide'}
#Ex-4
summary(titanic)
```
5. Being sure to allow for missing values if necessary, calculate (or plot) and interpret the following:
  - The standard deviation of fares
  - The 90th Percentile of fares 
  - Histogram of fares
```{r echo=FALSE,results='hide',fig.show='hide'}
#Ex-5
sd(titanic$fare, na.rm = TRUE)
quantile(titanic$fare, 0.9, na.rm = TRUE)
plot_ly(data = titanic, x = fare, type = "histogram")
```
6. Create a boxplot of ``fare`` by ``pclass`` and interpret your results.
```{r echo=FALSE,results='hide',fig.show='hide'}
#Ex-6
plot_ly(data = titanic, y = fare, group = pclass, type = "box", boxpoints = FALSE)
```
7. Create a data table called survivors, containing data for only those 
passengers who survived the disaster and answer the following questions:
  - What was the average age of the survivors? 
  - Among the survivors, were there more men or women? 
  - Do you notice anything different about social class for survivors compared to all passengers? 
```{r echo=FALSE,results='hide',fig.show='hide'}
#Ex-7
survivors <- titanic[survived == 1]
summary(survivors)
```
8. Determine whether men or women paid more, on average, for 
passage on the titanic. Hint: use something that looks like the following
``dataset[, operation, by = group]``
```{r echo=FALSE,results='hide',fig.show='hide'}
#Ex-8
titanic[, mean(fare, na.rm = TRUE), by = sex]
```
9. How did the fraction of survivors vary by ``sex`` ``pclass``? Can you think of a possible explanation for the pattern you see in the data? Hint: use something that looks like the following
``dataset[, operation, by = list(group1, group2)]``
```{r echo=FALSE,results='hide',fig.show='hide'}
#Ex-10
titanic[, mean(survived, na.rm = TRUE), by = list(sex, pclass)]
```
